---
title: "Using cmdstanr with NMI: Modern Bayesian Inference"
author: "NMI Package Authors"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Using cmdstanr with NMI: Modern Bayesian Inference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  eval = FALSE
)
```

## Introduction

The **nmi** package uses **cmdstanr** as its Bayesian inference backend, providing significant advantages over traditional approaches. This vignette explains the cmdstanr integration and how to optimize your NMI analyses.

## Why cmdstanr?

### Advantages of cmdstanr

- **⚡ Performance**: 2-3x faster sampling than rstan
- **🔧 Better diagnostics**: Enhanced R̂, ESS, and divergent transition warnings
- **🔄 Parallel processing**: Automatic multi-core chain execution
- **💾 Memory efficiency**: Lower memory usage for large models
- **🛠 Modern toolchain**: Latest Stan features and improvements

### From rstan to cmdstanr

The NMI package has been updated to use cmdstanr while maintaining backward compatibility with existing workflows.

## Installation and Setup

### Installing cmdstanr

```{r install-cmdstanr, eval=FALSE}
# Install cmdstanr from Stan repository
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))

# Load and install CmdStan backend
library(cmdstanr)
install_cmdstan()
```

### Verifying Installation

```{r verify-installation}
library(cmdstanr)

# Check CmdStan version
cmdstan_version()

# Check toolchain
check_cmdstan_toolchain()
```

### Troubleshooting Installation

```{r troubleshooting}
# If you encounter issues, try:
install_cmdstan(cores = 2)  # Use fewer cores
install_cmdstan(overwrite = TRUE)  # Force reinstall

# Check installation path
cmdstan_path()
```

## NMI with cmdstanr

### Automatic Integration

The NMI package automatically handles cmdstanr integration:

```{r auto-integration}
library(nmi)

# cmdstanr is automatically used - no additional setup needed!
nmi_results <- nmi_full_analysis(
  IPD = Example_IPD,
  AgD = Example_AgD_NMI,
  x_vect = c(0.6, 0.4),
  AgD_EM_cols = c("x1", "x2"),
  IPD_EM_cols = c("x1", "x2"),
  IPD_treatment_col = "Tr",
  AgD_treatment_cols = c("Trt1", "Trt2"),
  IPD_outcome_col = "Y",
  AgD_TE_col = "TE",
  AgD_SE_col = "se",
  AgD_study_col = "Study",
  study_sample_sizes = rep(300, 6),
  outcome_type = "binary"
)
```

### MCMC Settings

Control cmdstanr behavior through MCMC settings:

```{r mcmc-settings}
# Customize MCMC parameters
mcmc_settings <- list(
  n_iter = 2000,         # Total iterations per chain
  n_warmup = 1000,       # Warmup iterations
  n_chains = 4,          # Number of chains (parallel)
  adapt_delta = 0.95,    # Target acceptance rate
  max_treedepth = 12     # Maximum tree depth
)

# Use in analysis
nmi_results <- nmi_full_analysis(
  IPD = Example_IPD,
  AgD = Example_AgD_NMI,
  x_vect = c(0.6, 0.4),
  # ... other parameters ...
  mcmc_settings = mcmc_settings
)
```

### Performance Optimization

#### Parallel Chains

```{r parallel-chains}
# Automatic parallel execution (recommended)
mcmc_settings <- list(
  n_chains = 4,          # 4 chains run in parallel
  n_iter = 4000,         # More iterations for better estimates
  n_warmup = 2000
)
```

#### Memory Management

```{r memory-management}
# For large datasets, consider:
mcmc_settings <- list(
  n_chains = 2,          # Fewer chains for large models
  n_iter = 2000,         # Sufficient for most analyses
  adapt_delta = 0.8      # Standard setting
)
```

## Convergence Diagnostics

### Enhanced Diagnostics with cmdstanr

```{r enhanced-diagnostics}
# Check convergence automatically
nmi_results <- nmi_full_analysis(...)

# Extract diagnostics
if (!is.null(nmi_results$nma_results$BUGSoutput)) {
  summary_data <- nmi_results$nma_results$BUGSoutput$summary
  
  # R-hat values (should be < 1.1)
  rhat_values <- summary_data[, "Rhat"]
  max_rhat <- max(rhat_values, na.rm = TRUE)
  
  # Effective sample size
  ess_values <- summary_data[, "n.eff"]
  min_ess <- min(ess_values, na.rm = TRUE)
  
  cat("Maximum R̂:", round(max_rhat, 3), "\n")
  cat("Minimum ESS:", round(min_ess, 0), "\n")
  
  # Convergence assessment
  if (max_rhat < 1.1 && min_ess > 100) {
    cat("✓ Good convergence\n")
  } else {
    cat("⚠ Check convergence\n")
  }
}
```

### Detailed Convergence Checking

```{r detailed-convergence}
# Access the cmdstanr fit object directly
cmdstan_fit <- nmi_results$nma_results$cmdstan_fit

# Check sampling diagnostics
cmdstan_fit$diagnostic_summary()

# Check for divergent transitions
sampler_diagnostics <- cmdstan_fit$sampler_diagnostics()
divergent_transitions <- sum(sampler_diagnostics[,,"divergent__"])
cat("Divergent transitions:", divergent_transitions, "\n")
```

## Advanced Usage

### Custom Stan Models

You can also use the Stan models directly:

```{r custom-models}
# Get the Stan model code
model_code <- stan_nma_model("binary")
cat(model_code)

# Compile and use manually
library(cmdstanr)
model <- cmdstan_model(write_stan_file(model_code))

# Prepare data
stan_data <- list(
  ns = 10,           # number of studies
  nt = 3,            # number of treatments  
  t = matrix(c(1,2), nrow=10, ncol=2),  # treatment indicators
  y = rnorm(10),     # outcomes
  se = rep(0.1, 10)  # standard errors
)

# Sample
fit <- model$sample(data = stan_data, chains = 2)
```

### Multiple Outcome Types

```{r multiple-outcomes}
# Binary outcomes
nmi_binary <- nmi_full_analysis(..., outcome_type = "binary")

# Continuous outcomes  
nmi_continuous <- nmi_full_analysis(..., outcome_type = "continuous")

# Count outcomes
nmi_count <- nmi_full_analysis(..., outcome_type = "count")
```

## Troubleshooting

### Common Issues

#### CmdStan Not Found

```{r cmdstan-not-found}
# If you see "CmdStan not found"
library(cmdstanr)
install_cmdstan()

# Or specify path manually
set_cmdstan_path("/path/to/cmdstan")
```

#### Compilation Errors

```{r compilation-errors}
# Check C++ toolchain
check_cmdstan_toolchain()

# Update cmdstanr
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
```

#### Convergence Issues

```{r convergence-issues}
# If chains don't converge, try:
mcmc_settings <- list(
  n_iter = 4000,         # More iterations
  n_warmup = 2000,       # Longer warmup
  adapt_delta = 0.95,    # Higher target acceptance rate
  max_treedepth = 12     # Deeper trees
)
```

### Performance Tips

#### For Large Datasets

```{r large-datasets}
# Reduce memory usage
mcmc_settings <- list(
  n_chains = 2,          # Fewer chains
  n_iter = 1000,         # Fewer iterations initially
  n_warmup = 500
)
```

#### For Complex Models

```{r complex-models}
# Increase adaptation
mcmc_settings <- list(
  n_iter = 3000,
  n_warmup = 1500,       # 50% warmup
  adapt_delta = 0.9,     # Conservative sampling
  max_treedepth = 12
)
```

## Comparison with rstan

### Performance Comparison

| Feature | rstan | cmdstanr |
|---------|-------|----------|
| Speed | 1x | 2-3x faster |
| Memory | Higher | Lower |
| Diagnostics | Basic | Enhanced |
| Parallel | Manual | Automatic |
| Stan Version | Older | Latest |

### Migration Benefits

- **Automatic**: No code changes needed for basic usage
- **Backward Compatible**: Existing workflows still work
- **Enhanced Output**: Better diagnostics and warnings
- **Future-Proof**: Active development and support

## Best Practices

### MCMC Settings

```{r best-practices-mcmc}
# Recommended settings for most analyses
recommended_settings <- list(
  n_iter = 2000,         # Good balance of speed/accuracy
  n_warmup = 1000,       # 50% warmup
  n_chains = 4,          # Parallel chains for diagnostics
  adapt_delta = 0.8,     # Default, increase if divergences
  max_treedepth = 10     # Default, increase if warnings
)
```

### Convergence Monitoring

```{r best-practices-convergence}
# Always check convergence
check_convergence <- function(nmi_results) {
  if (!is.null(nmi_results$nma_results$BUGSoutput)) {
    summary_data <- nmi_results$nma_results$BUGSoutput$summary
    max_rhat <- max(summary_data[, "Rhat"], na.rm = TRUE)
    min_ess <- min(summary_data[, "n.eff"], na.rm = TRUE)
    
    return(list(
      converged = max_rhat < 1.1 && min_ess > 100,
      max_rhat = max_rhat,
      min_ess = min_ess
    ))
  }
  return(NULL)
}
```

## References

- **cmdstanr documentation**: https://mc-stan.org/cmdstanr/
- **Stan User's Guide**: https://mc-stan.org/docs/
- **NMI Methodology**: Harari et al. (2023)

## Session Info

```{r session-info}
sessionInfo()
``` 