---
title: "Continuous Effect Modifiers in NMI"
author: "Ahmad Sofi-Mahmudi"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Continuous Effect Modifiers in NMI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
is_check <- ("CheckExEnv" %in% search()) || any(c("_R_CHECK_TIMINGS_", "_R_CHECK_LICENSE_") %in% names(Sys.getenv()))
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  eval = !is_check
)
```

# Introduction

The Network Meta-Interpolation (NMI) package has been extended to support **continuous effect modifiers**, moving beyond the original binary-only implementation. This advancement allows researchers to handle real-valued covariates like age, BMI, or biomarker levels in network meta-analysis.

## Installation

If you haven't installed the NMI package yet:

```{r installation, eval=FALSE}
# Install from GitHub (recommended)
devtools::install_github("choxos/nmi", build_vignettes = TRUE)

# Or install locally if you have the source
# devtools::install_local("path/to/nmi", build_vignettes = TRUE)

# For Bayesian features (optional):
# install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
# cmdstanr::install_cmdstan()
# cmdstanr::check_cmdstan_toolchain(fix = TRUE)

# Load the package
library(nmi)
```

## What's New in v1.1.0+

- 🎯 **Continuous Effect Modifier Support**: Handle age, BMI, continuous biomarkers
- 📈 **Multiple Interpolation Methods**: Linear, spline, and adaptive discretization
- 🔍 **Automatic EM Type Detection**: Smart detection of binary vs continuous variables
- ⚡ **Flexible API**: Seamless integration with existing NMI workflow
- ✅ **Comprehensive Testing**: Full validation suite with real-world examples

## Key Advantages

1. **No Information Loss**: Work directly with continuous data without forced discretization
2. **Flexible Modeling**: Choose between linear, spline, or optimal discretization approaches
3. **Robust Extrapolation Warnings**: Automatic detection of extrapolation scenarios
4. **Future-Ready**: Foundation for mixed EM types and multivariate methods

---

# Core Concepts

## Effect Modifier Types

The NMI package now automatically detects three types of effect modifiers:

- **Binary**: Two values (0,1) - traditional NMI approach
- **Categorical**: Few discrete integer values (1,2,3,...)  
- **Continuous**: Many unique values or non-integer data

## Interpolation Methods

### 1. Linear Interpolation
- **Best for**: Smooth, approximately linear relationships
- **Advantages**: Simple, robust, fast
- **Use when**: Limited data points or linear EM-effect relationship

### 2. Spline Interpolation  
- **Best for**: Non-linear, smooth relationships
- **Types**: Natural splines, cubic splines, smoothing splines
- **Use when**: Sufficient data points (≥3) and non-linear relationships

### 3. Adaptive Discretization
- **Best for**: Hybrid approach maintaining interpretability
- **Methods**: Equal frequency, equal width, outcome-optimized
- **Use when**: Need to balance continuous modeling with discrete analysis

---

# Getting Started

## Basic Example: Age as Continuous EM

```{r basic_example}
library(nmi)

# Create example data with age as continuous effect modifier
set.seed(2025)

# Individual Patient Data
IPD <- data.frame(
  Study = rep("IPD_Study", 100),
  age = rnorm(100, mean = 60, sd = 10),
  sex = rbinom(100, 1, 0.5),
  treatment = sample(c("A", "B"), 100, replace = TRUE),
  outcome = rbinom(100, 1, 0.4)
)

# Aggregate Data with different age profiles
AgD <- data.frame(
  Study = paste0("Study_", 1:5),
  age_mean = c(45, 55, 65, 70, 75),
  TE = c(0.1, 0.3, 0.5, 0.4, 0.2),
  se = c(0.12, 0.10, 0.11, 0.13, 0.15)
)

print(head(IPD))
print(AgD)
```

## Automatic EM Type Detection

```{r em_detection}
# Detect effect modifier types automatically
em_types <- detect_em_types(IPD, c("age", "sex"))
print(em_types)
```

The function correctly identifies:
- `age` as continuous (many unique values)
- `sex` as binary (only 0 and 1 values)

---

# Interpolation Methods

## Linear Interpolation

Linear interpolation provides a simple and robust method for continuous effect modifiers:

```{r linear_interpolation}
# Target age for interpolation
target_age <- 62.5

# Perform linear interpolation
linear_result <- linear_interpolation(
  study_data = AgD,
  target_values = c(age_mean = target_age),
  em_cols = "age_mean",
  te_col = "TE",
  se_col = "se"
)

print(linear_result)
```

### Extrapolation Warnings

The function automatically detects and warns about extrapolation:

```{r linear_extrapolation}
# Extrapolation example (age outside observed range)
extrap_result <- linear_interpolation(
  study_data = AgD,
  target_values = c(age_mean = 85),  # Outside range
  em_cols = "age_mean"
)

print(paste("Extrapolation detected:", extrap_result$extrapolation))
```

## Spline Interpolation

For non-linear relationships, spline interpolation offers greater flexibility:

```{r spline_interpolation}
# Natural spline interpolation
spline_result <- spline_interpolation(
  study_data = AgD,
  target_values = c(age_mean = target_age),
  em_cols = "age_mean",
  spline_type = "natural"
)

print(spline_result)

# Compare methods
comparison <- data.frame(
  Method = c("Linear", "Spline"),
  TE = c(linear_result$te, spline_result$te),
  SE = c(linear_result$se, spline_result$se)
)
print(comparison)
```

### Smoothing Splines

For noisy data, smoothing splines can provide better results:

```{r smoothing_spline}
# Smoothing spline (handles noise better)
smooth_result <- spline_interpolation(
  study_data = AgD,
  target_values = c(age_mean = target_age),
  em_cols = "age_mean",
  spline_type = "smoothing"
)

print(smooth_result)
```

---

# Adaptive Discretization

When continuous modeling isn't suitable, adaptive discretization finds optimal binning:

```{r discretization}
# Generate continuous age data
age_data <- rnorm(200, 60, 12)

# Optimize discretization
disc_result <- optimize_discretization(
  continuous_data = age_data,
  method = "equal_freq",
  max_bins = 4
)

print(disc_result$bin_counts)
print(paste("Optimal bins:", disc_result$n_bins))
print(paste("Optimization score:", round(disc_result$score, 3)))
```

## Discretization Methods

```{r discretization_methods}
# Compare discretization methods
freq_disc <- optimize_discretization(age_data, method = "equal_freq", max_bins = 3)
width_disc <- optimize_discretization(age_data, method = "equal_width", max_bins = 3)

cat("Equal Frequency bins:", freq_disc$n_bins, "\n")
cat("Equal Width bins:", width_disc$n_bins, "\n")
```

---

# Full NMI Analysis with Continuous EMs

The main NMI function now supports continuous effect modifiers:

```{r full_nmi_analysis}
# Full continuous NMI analysis
target_values <- c(age = 67.5)

nmi_result <- NMI_interpolation_continuous(
  IPD = IPD,
  AgD = AgD,
  x_vect = target_values,
  AgD_EM_cols = "age_mean",
  IPD_EM_cols = "age",
  interpolation_method = "linear"
)

print(nmi_result$Final)
print(paste("Interpolation method:", nmi_result$method))
print(paste("Continuous EMs:", nmi_result$continuous_ems))
```

---

# Visualization

Visualizing the age-treatment effect relationship helps validate interpolation:

```{r visualization, fig.width=8, fig.height=6}
# Create age sequence for visualization
age_seq <- seq(40, 80, by = 2.5)

# Interpolate treatment effects across age range
interpolated_effects <- sapply(age_seq, function(age) {
  tryCatch({
    result <- linear_interpolation(AgD, c(age_mean = age), "age_mean")
    result$te
  }, error = function(e) NA)
})

# Create plot data
plot_data <- data.frame(
  age = age_seq,
  treatment_effect = interpolated_effects
)

# Remove NA values (from extrapolation failures)
plot_data <- plot_data[!is.na(plot_data$treatment_effect), ]

# Plot if ggplot2 is available
if (requireNamespace("ggplot2", quietly = TRUE)) {
  library(ggplot2)
  
  p <- ggplot() +
    geom_line(data = plot_data, 
              aes(x = age, y = treatment_effect), 
              color = "blue", size = 1) +
    geom_point(data = AgD, 
               aes(x = age_mean, y = TE), 
               color = "red", size = 3) +
    geom_vline(xintercept = target_values["age"], 
               linetype = "dashed", color = "green") +
    labs(
      title = "Age-Treatment Effect Relationship",
      subtitle = "Blue: Interpolated effects, Red: Observed data, Green: Target age",
      x = "Age (years)",
      y = "Treatment Effect"
    ) +
    theme_minimal()
  
  print(p)
} else {
  # Alternative base R plot
  plot(AgD$age_mean, AgD$TE, 
       pch = 19, col = "red", cex = 1.5,
       xlab = "Age (years)", ylab = "Treatment Effect",
       main = "Age-Treatment Effect Relationship")
  lines(plot_data$age, plot_data$treatment_effect, 
        col = "blue", lwd = 2)
  abline(v = target_values["age"], 
         lty = 2, col = "green", lwd = 2)
  legend("topright", 
         c("Observed", "Interpolated", "Target"), 
         col = c("red", "blue", "green"), 
         pch = c(19, NA, NA), lty = c(NA, 1, 2))
}
```

---

# Advanced Features

## Method Comparison

Compare different interpolation methods across multiple target values:

```{r method_comparison}
# Compare methods across age range
comparison_ages <- c(50, 60, 65, 70)
methods_comparison <- data.frame(
  age = comparison_ages,
  linear_te = NA,
  spline_te = NA
)

for (i in seq_along(comparison_ages)) {
  age <- comparison_ages[i]
  
  # Linear interpolation
  linear_res <- linear_interpolation(AgD, c(age_mean = age), "age_mean")
  methods_comparison$linear_te[i] <- linear_res$te
  
  # Spline interpolation
  spline_res <- spline_interpolation(AgD, c(age_mean = age), "age_mean")
  methods_comparison$spline_te[i] <- spline_res$te
}

print(round(methods_comparison, 3))
```

## Error Handling

The package provides comprehensive error handling and informative messages:

```{r error_handling}
# Example: Multiple continuous EMs (not yet supported)
tryCatch({
  NMI_interpolation_continuous(
    IPD = IPD,
    AgD = AgD,
    x_vect = c(age = 60, bmi = 25),
    AgD_EM_cols = c("age_mean", "bmi_mean"),
    IPD_EM_cols = c("age", "bmi"),
    em_types = c(age = "continuous", bmi = "continuous")
  )
}, error = function(e) {
  cat("Expected error:", e$message, "\n")
})
```

---

# Best Practices

## When to Use Each Method

### Linear Interpolation
- ✅ **Use when**: Linear relationship between EM and treatment effect
- ✅ **Use when**: Limited data points (< 5 studies)
- ✅ **Use when**: Need simple, interpretable results
- ❌ **Avoid when**: Clear non-linear relationships exist

### Spline Interpolation  
- ✅ **Use when**: Non-linear relationships expected
- ✅ **Use when**: Sufficient data points (≥ 5 studies)
- ✅ **Use when**: Smooth underlying relationship assumed
- ❌ **Avoid when**: Very noisy data or few observations

### Adaptive Discretization
- ✅ **Use when**: Need to maintain compatibility with binary NMI
- ✅ **Use when**: Interpretability is crucial
- ✅ **Use when**: Continuous modeling fails or is inappropriate
- ❌ **Avoid when**: Information loss is a major concern

## Data Requirements

### Minimum Requirements
- **Linear**: ≥ 2 data points
- **Spline**: ≥ 3 data points (falls back to linear if fewer)
- **Discretization**: ≥ 10 observations for stable binning

### Recommended
- **Linear**: 3-5 data points across EM range
- **Spline**: 5-10 data points for stable estimation
- **Discretization**: ≥ 50 observations for optimal binning

---

# Limitations and Future Work

## Current Limitations

1. **Single Continuous EM**: Only one continuous EM supported per analysis
2. **No Mixed Types**: Cannot combine binary and continuous EMs yet
3. **Limited Multivariate**: No multivariate spline methods implemented
4. **Simple Discretization**: Advanced discretization algorithms planned

## Planned Enhancements (v1.2.0+)

1. **Multiple Continuous EMs**: Support for multivariate continuous modeling
2. **Mixed EM Types**: Combine binary, categorical, and continuous EMs
3. **Advanced Splines**: Tensor product splines, adaptive knot selection
4. **ML-Based Discretization**: Outcome-optimized binning algorithms
5. **Uncertainty Quantification**: Bootstrap confidence intervals for interpolation

---

# Technical Details

## Algorithm Overview

The continuous EM implementation follows this workflow:

1. **EM Type Detection**: Automatic classification of variable types
2. **Data Validation**: Check for sufficient observations and valid ranges
3. **Method Selection**: Choose appropriate interpolation approach
4. **Interpolation**: Apply selected method with error handling
5. **Extrapolation Warning**: Detect and warn about out-of-range predictions
6. **Result Integration**: Format results for compatibility with existing NMI

## Performance Considerations

- **Linear Interpolation**: O(n log n) complexity (sorting)
- **Spline Interpolation**: O(n²) for smoothing splines
- **Discretization**: O(n²k) where k is maximum bins considered

For large datasets (>1000 studies), consider using linear interpolation or pre-filtering to most relevant studies.

---

# Session Information

```{r session_info}
sessionInfo()
```

---

# References

1. Harari et al. (2023). Network meta-interpolation: Effect modification adjustment in network meta-analysis using subgroup analyses.

2. Sofi-Mahmudi, A. (2025). NMI Package: Continuous Effect Modifier Extensions. R package version 1.1.0+.

3. Wood, S.N. (2017). Generalized Additive Models: An Introduction with R (2nd edition). Chapman and Hall/CRC.

For more information, visit the [NMI package repository](https://github.com/choxos/nmi) or contact Ahmad Sofi-Mahmudi at a.sofimahmudi@gmail.com. 