---
title: "Getting Started with NMI"
author: "Ahmad Sofi-Mahmudi"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Getting Started with NMI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
is_check <- ("CheckExEnv" %in% search()) || any(c("_R_CHECK_TIMINGS_", "_R_CHECK_LICENSE_") %in% names(Sys.getenv()))
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  eval = !is_check
)
```

# Introduction

Welcome to the Network Meta-Interpolation (NMI) package! This vignette will guide you through the basic concepts and help you get started with your first NMI analysis.

## What is Network Meta-Interpolation?

Network Meta-Interpolation (NMI) is a methodology for addressing effect modification in network meta-analysis when combining individual patient data (IPD) and aggregate data (AgD). The method allows researchers to:

- Perform network meta-analysis at specific covariate levels
- Handle effect modification in mixed IPD/AgD networks  
- Support multiple outcome types (binary, continuous, count)
- Work with various effect modifier types (binary, categorical, continuous)

## Installation

### From GitHub (Recommended)

```{r installation, eval=FALSE}
# Install the latest stable version
devtools::install_github("choxos/nmi")

# Install with vignettes (recommended for learning)
devtools::install_github("choxos/nmi", build_vignettes = TRUE)

# Install development version (latest features)
devtools::install_github("choxos/nmi", ref = "develop", build_vignettes = TRUE)
```

### Local Installation

If you have the source code locally:

```{r local_installation, eval=FALSE}
# Install from local directory
devtools::install_local("path/to/nmi", build_vignettes = TRUE)
```

### Load the Package

```{r load_package}
library(nmi)
```

## Package Overview

The NMI package provides a comprehensive suite of functions for network meta-analysis with effect modification support: 